### 2024-12-02 paper/paper.qmd
Wrote the modeling section.

---
title: "From Light to Growth: A Study of Night-Time Light as an Indicator of Economic Development"
subtitle: "Growth in Night-Time Light Intensity Correlates ___% With GDP Growth"
author: 
  - Shamayla Durrin
thanks: "Code and data are available at: [https://github.com/shamayla38/EconomyLightCorrelation](https://github.com/shamayla38/EconomyLightCorrelation)."
date: today
date-format: long
abstract: "This study investigates the relationship between the growth of night-time light (NTL) intensity, as captured by satellite imagery, and GDP growth across countries over a multi-decade period. By analyzing data spanning from 1992 to 2001, we find a strong correlation of [X]% between NTL growth and GDP growth. Importantly, the strength of this correlation varies by the statistical capacity of countries. Countries with high statistical capacity, rated 'A' by the World Bank, exhibit a correlation as high as [X]%, while those with low statistical capacity, rated 'F', show a correlation as low as [Y]%. These findings highlight the potential of NTL data as a complementary tool for economic analysis, particularly in regions where conventional economic metrics are less reliable or unavailable.(Blank spaces will be filled when project is completed)"
format: pdf
number-sections: true
bibliography: references.bib
execute:
  echo: false
  warning: false
  message: false
header-includes:
  - \usepackage[margin=0.8in]{geometry}
---

# Introduction

# Data

We collected data from multiple sources to analyze the relationship between night-time light intensity and GDP growth. The primary dataset is a harmonized global night-time light (NTL) dataset spanning from 1992 to 2018, developed by Li et al. [@Li2020] The dataset combines observations from two satellite systems: DMSP/OLS (1992–2013) and VIIRS (2012–2018). To ensure consistency across years, they addressed differences in spatial resolution and saturation between the two systems by inter-calibrating DMSP data and simulating DMSP-like VIIRS observations. The resulting dataset provides annual Digital Number (DN) values, representing night-time illumination intensity, as GeoTIFF files.

We also sourced GDP, manufacturing share of GDP, and population data from the World Bank’s World Development Indicators (WDI) [@WorldBankWDI],measuring GDP in purchasing power parity (PPP) for comparability across countries. Additionally, we incorporated the Statistical Performance Indicator (SPI), which evaluates national statistical systems across five pillars: data use, data services, data products, data sources, and data infrastructure. [@WorldBankSPI] A thorough discussion on how we organised the data set is given in [@sec-clean].

## Measurement

Night-time light intensity, which reflects human activity, is indirectly measured by satellite sensors onboard the DMSP/OLS and VIIRS systems. These satellites capture light emissions from streetlights, industrial complexes, and residential areas, converting this physical phenomenon into Digital Number (DN) values. The brightness levels are then transformed into quantifiable data, which are stored in GeoTIFF format for further analysis. DMSP/OLS, operational since the 1970s, used low-light imaging to detect artificial lights but had limitations such as coarse resolution and sensor saturation [@elvidge1997]. VIIRS, launched in 2012, provided better spatial resolution and could capture a wider range of light levels, with improved algorithms to filter out unwanted noise and natural light sources.[@elvidge2017]. The harmonized version of these two datasets, combining data from DMSP/OLS and VIIRS, has been used by researchers for analysis.

GDP is measured as the total economic output of a country, adjusted for purchasing power parity (PPP) to standardize across different national economies. This transformation from an elaborate economic reality into a single numerical value is done through extensive national accounting systems, with data collected by governments through economic surveys, reports, and censuses.Population figures, reflecting demographic changes and the size of human capital, are also derived from national censuses and are measured in absolute numbers. Meanwhile, the manufacturing share of GDP is calculated by extracting relevant economic data, representing the portion of a country's GDP generated by the manufacturing sector. The World Bank sources all these data from national accounts.

## Data Extraction and Cleaning {#sec-clean}

The data preparation began with extracting night-time light (NTL) intensity values from GeoTIFF files using Python's 'Rasterio' [@raster] package. This process involved reading annual TIFF files to obtain Digital Number (DN) values, representing light intensity, along with their corresponding latitude and longitude coordinates. This resulted in a spatial dataset detailing global NTL distribution for each year.

To associate each observation with a specific country, the extracted data were merged with a global shapefile of country boundaries through spatial join using [@geopandas]. Each latitude-longitude pair was assigned to the country in which it was located, and observations without an assigned country—primarily from international waters—were removed. Subsequently, to create an aggregate annual measure of NTL intensity for each country, all observations within a country were summed to calculate the total light emitted annually.

The World Bank data, including GDP, manufacturing share of GDP, and population, were processed by standardizing country names to ensure consistency across datasets. For instance, discrepancies like 'Yemen' versus 'Yemen, Rep.' were resolved. The data were then transformed from a wide to a long format using pandas' 'melt' function, facilitating easier merging with the NTL data.

To assess how the correlation between night-time light (NTL) intensity and GDP varies among countries with differing levels of statistical capacity, we first to group countries together based on a measure of each country's statistical capacity.We used the World Bank's Statistical Performance Indicators (SPI). We calculated the average SPI score for each country across all the years SPI was available for and assigned grades based on these averages: scores of 80 and above received an 'A', scores between 60 and 79 a 'B', scores between 40 and 59 a 'C', scores between 20 and 39 a 'D', and scores below 20 an 'F'.A glimpse of the dataset is presented in [@tbl-glimpse].

```{r}

#| echo: false
#| warning: false
#| message: false

library(arrow)        
library(tidyverse)         
library(modelsummary)
library(here)
library(kableExtra)
library(tinytable)
library(sf)
library(scales)
library(gridExtra)
library(grid)
library(png)
library(modelsummary)
library(broom)

# Load the Parquet file
analysis_data <- read_parquet(here("data/02-analysis_data/04-analysis/analysis.parquet"))

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-glimpse
#| tbl-cap: "Sample of the dataset showing key variables used in the analysis"

# Filter rows without missing values, pick the first occurrence of 10 unique countries
clean_data <- analysis_data[complete.cases(analysis_data), ] |>  # Remove rows with missing values
  group_by(country) |>  # Group by country
  slice_head(n = 1) |>  # Select the first row for each country
  ungroup() |>  # Remove grouping
  slice_head(n = 7)  # Limit to 10 unique countries

# Rename columns using base R
colnames(clean_data) <- c(
  "Country", 
  "Year", 
  "Night Lights (DN)", 
  "GDP (PPP)", 
  "Manufacturing Share", 
  "Country Code", 
  "Population", 
  "SPI Score", 
  "Grade"
)

# Drop the "SPI Score" and "Country Code" columns
clean_data <- clean_data[, !colnames(clean_data) %in% c("SPI Score", "Country Code")]

# Display the cleaned data as a tidy table
clean_data |> tt()

```

## Summary Statistics

It is more meaningful to examine the summary statistics year-wise to capture temporal variations in the data, as the variance across all years is substantial. [@tbl-sumstat] displays the summary statistics for selected years, showing significant growth in average night-time light intensity (DN), GDP (measured in PPP), and population from 1995 to 2020. An especially intriguing observation is that both night-time light intensity and GDP grew by almost exactly the same proportion—2.1 times—between 1995 and 2020 (DN increased by 2.102 times and GDP by 2.101 times).The summary statistics of the rest of the years are given in appendix.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-sumstat
#| tbl-cap: "Summary Statistics of Data from year 1995 and 2020"

# Filter the dataset for selected years and select relevant columns
filtered_data <- analysis_data %>%
  filter(year %in% c(1995,2020)) %>%  # Filter for specific years
  select(
    "Year" = year,
    "DN" = dn,
    "GDP" = gdp,
    "Manufacturing" = manufacturingsharegdp,
    "Population" = population
  )

# Generate descriptive statistics by year
datasummary_balance(
  formula = ~Year,  # Group by 'Year'
  data = filtered_data,
  dinm = FALSE,  # Prevent dynamic naming
  notes = "DN refers to total average night-time light intensity emitted by a country in a year, GDP is measured in PPP, and Manufacturing refers to the manufacturing share of GDP."
)

```

```{r}
#| echo: false
#| warning: false
#| message: false
# Load the shapefile and suppress messages
shapefile_path <- "data/01-raw_data/02-shapefile"
sink("/dev/null")  # Redirect console output to null
world_shapefile <- st_read(here(shapefile_path))
sink()# Load the Parquet file
metrics <- read_parquet(here("data/03-plotting_data/plotting.parquet"))
```

## Data Description

### Night Time Lights Data

Nighttime light (NTL) distribution and intensity often vary significantly by population density, GDP, and levels of urbanization. When comparing countries with similar population sizes and land areas, the disparity in NTL intensity vividly reflects income inequality. A striking example is the contrast between North and South Korea, where South Korea's GDP per capita is over 50 times greater than that of North Korea, manifesting in the stark difference in nighttime illumination. In [@fig-ntl], using NASA's Worldview NTL imagery, this difference of GDP is clearly visible, with South Korea brightly lit compared to its nearly dark northern neighbor. Similarly, Myanmar and Thailand, with comparable populations and land area, show a distinct NTL difference; Thailand's higher GDP and urbanization are evident in its dense and widespread lighting compared to Myanmar's relatively sparse illumination.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ntl
#| fig-cap: "Night-time light intensity highlights stark contrasts in economic development: the profound disparity between North and South Korea, and the significant difference between Myanmar and Thailand.*(Source: NASA Worldview)*"

knitr::include_graphics(here("others/plots/ntl.png"))

```

Total night-time light (NTL) values for countries exhibit a highly skewed distribution due to the presence of a few highly luminous regions. In [@fig-dn], we observe that after applying a log transformation, this distribution becomes relatively normal, making it more suitable for comparative and statistical analysis

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dn
#| fig-cap: "Distribution of log-transformed average nighttime luminosity (DN) by country over the past 30 years, showing variations in average brightness levels across nations."

# Group by country and calculate the average DN for each country
avg_dn_by_country <- analysis_data %>%
  group_by(country) %>%
  summarise(avg_dn = mean(dn, na.rm = TRUE)) %>%
  ungroup()

ggplot(avg_dn_by_country, aes(x = avg_dn)) +
  geom_histogram(fill = "steelblue", color = "white", alpha = 0.7, bins = 30) +
  scale_x_continuous(trans = "log10", labels = scales::comma_format()) +
  labs(
    x = "Average DN (Log Scale)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

### GDP

To visualise GDP distribution across numerous countries, we categorized them into UN-defined SDG regions and examined the GDP distribution for 2012-2020. The [@fig-gdp] shows that Europe and Northern America have the highest GDP levels, followed by Eastern and South-Eastern Asia and Latin America and the Caribbean, while Sub-Saharan Africa, Oceania, and Central and Southern Asia show significantly lower GDP levels. We would expect similar variations in night-time light intensity, with regions of higher GDP likely exhibiting brighter and more widespread night-time illumination. This connection underscores how night-time light can serve as a proxy for economic activity and development, reflecting the stark differences across regions.

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-gdp
#| fig-cap: "Density plot of average GDP (2012-2020) by region, illustrating Europe and Northern America with the highest GDP, followed by Eastern and South-Eastern Asia and Latin America and the Caribbean, while Sub-Saharan Africa, Oceania, and Central and Southern Asia exhibit significantly lower GDP levels."

# Filter data for the speicific interval of years 
filtered_data_first_interval <- metrics %>%
  filter(interval == "2012-2020", !is.na(avg_gdp), !is.na(region))

# Create a density plot for the first interval
ggplot(filtered_data_first_interval, aes(x = avg_gdp, fill = region)) +
  geom_density(alpha = 0.6) +
  scale_x_continuous(trans = "log10", 
                     labels = scales::comma_format()) +  # Log scale for better visualization
  labs(
    x = "Average GDP (log scale)",
    y = "Density",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

```

### Manufacturing Share of GDP of Country

Night-time light intensity is expected to vary with the share of manufacturing in a country's GDP, as higher manufacturing activity typically involves more factories, industrial areas, and infrastructure requiring artificial lighting. [@fig-man] in appendix illustrates the manufacturing share of GDP for countries in 2020. Most countries have a manufacturing share below 20%, with a few exceptions such as Puerto Rico, Liechtenstein, and Ireland, which have some of the highest manufacturing contributions. [@fig-man2] in the appendix presents the manufacturing share of GDP at the beginning of our dataset's time frame in 1992, highlighting how many countries have gradually shifted towards a higher reliance on manufacturing over the years.

### Population

Population trends provide essential context for understanding economic activity and night-time light (NTL) intensity patterns. In [@fig-pop], the total world population is shown to have grown steadily from approximately 4 billion in 1992 to over 7.5 billion by 2020. [@fig-pop2] provides a regional breakdown of average population across intervals, categorized by UN-defined SDG regions. It reveals that regions like Central and Southern Asia and Eastern and South-Eastern Asia consistently have the highest average populations, while regions like Oceania and Sub-Saharan Africa have significantly lower averages. This demographic variation is critical for analyzing NTL because densely populated regions are likely to exhibit higher light intensity due to concentrated urbanization and industrial activity. Conversely, less populated or sparsely developed regions are expected to have lower NTL levels.

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pop
#| fig-cap: "Steady growth of the global population from 1992 to 2020"

# Summarize population by year
total_population_by_year <- analysis_data %>%
  group_by(year) %>%
  summarise(total_population = sum(population, na.rm = TRUE))

# Create a bar chart
ggplot(total_population_by_year, aes(x = year, y = total_population)) +
  geom_bar(stat = "identity", fill = "blue", alpha = 0.7) +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(
    x = "Year",
    y = "Total World Population"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pop2
#| fig-cap: "Average population (log scale) across regions for different time intervals (1991–1997, 1998–2004, 2005–2011, and 2012–2020) showing significant population growth in regions such as Central and Southern Asia, Eastern and South-Eastern Asia, and Sub-Saharan Africa over time, with variations across other regions"

region_population <- metrics %>%
  filter(!is.na(region) & !is.na(avg_population)) %>%  # Remove rows with NA values in region or avg_population
  group_by(region, interval) %>%
  summarize(avg_population = mean(avg_population, na.rm = TRUE))

# Create a grouped bar plot
ggplot(region_population, aes(x = region, y = avg_population, fill = interval)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(
    x = "Region",
    y = "Average Population",
    fill = "Interval"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```
# Night lights and GDP

[@fig-log-dn-vs-log-gdp] plots the log of DN values against the log of GDP for all countries across 30 years, capturing the relationship between night-time light intensity and economic activity. The scatterplot shows a clear positive correlation, indicating that higher GDP values are generally associated with higher DN values.

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-log-dn-vs-log-gdp
#| fig-cap: "Strong positive correlation between night-time light intensity (Log DN) and GDP (Log GDP), indicating regions with higher GDP levels tend to emit more night-time light"

# Ensure data is cleaned (remove NA values for dn and gdp)
analysis_cleaned <- analysis_data %>%
  filter(!is.na(dn) & !is.na(gdp))

# Create the scatterplot
ggplot(analysis_cleaned, aes(x = log(gdp), y = log(dn))) +
   geom_point(alpha = 0.7, color = "gold4", size = 1) +
  labs(
    x = "Log of GDP",
    y = "Log of DN"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

```
In [@fig-dn-gdp-grade], we observe how this relationship between DN and GDP varies across countries with different statistical capacities. A list of countries categorized by their statistical capacity grades is given in [@tab-grades-countries-latex] of the appendix. Countries with higher statistical capacity (graded A and B) exhibit a more tightly clustered and linear relationship, indicating consistent and robust data collection processes. Conversely, countries with lower statistical capacity (graded D and F) show more scatter and variability, suggesting potential discrepancies or weaker data reliability in their national accounts. This shows the intuitive idea that the strength of the DN-GDP relationship can serve as an indicator of the reliability of a country's national statistical systems.

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dn-gdp-grade
#| fig-cap: "Strength of the correlation between Log DN and Log GDP varies with statistical capacity grades; countries with higher grades (A, B, C) exhibit stronger and more consistent relationships, while lower grades (D, F) show weaker and more dispersed patterns, suggesting the quality of national accounts impacts the NTL-GDP relationship."

# Ensure data is cleaned (remove NA values for dn, gdp, and grade)
analysis_cleaned <- analysis_data %>%
  filter(!is.na(dn) & !is.na(gdp) & !is.na(grade))

# Create the scatterplot faceted by grade
ggplot(analysis_cleaned, aes(x = log(gdp), y = log(dn))) +
  geom_point(alpha = 0.7, color = "gold4", size = 0.5) +
  labs(
    x = "Log of GDP",
    y = "Log of DN"
  ) +
  facet_wrap(~grade) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(face = "bold")
  )

```
# Night lights and Population

Larger populations naturally result in higher night-time light emissions due to increased urbanization and economic activity. Figure 1 confirms this with a clear positive relationship between population (log-transformed) and night-time light intensity (Log DN), highlighting how population size contributes to observable brightness from space.

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-population-vs-dn
#| fig-cap: "Scatterplot of Population versus DN."

# Ensure data is cleaned (remove NA values for population and dn)
analysis_cleaned <- analysis_data %>%
  filter(!is.na(population) & !is.na(dn))

# Create the scatterplot
ggplot(analysis_cleaned, aes(x = log(population), y = log(dn))) +
  geom_point(alpha = 0.7, color = "darkcyan", size = 0.5) +
  labs(
    x = "Log of Population",
    y = "Log of DN"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

# NTL and ith Manufacturing Share of GDP

[@fig-manufacturing-share-vs-dn] shows the relationship between the manufacturing share of GDP and night-time light intensity (Log DN). The plot shows a weak positive trend for countries with moderate manufacturing shares (up to 20%), where higher manufacturing activity corresponds to slightly higher DN values. However, the relationship becomes less defined for countries with very high or very low manufacturing shares, suggesting that other economic activities or structural factors might influence night-time light emissions. This indicates that while manufacturing contributes to night-time light, it is not the sole determinant, and its impact may vary across different contexts.

```{r fig.width=12, fig.height=5}
#| echo: false
#| warning: false
#| message: false
#| label: fig-manufacturing-share-vs-dn
#| fig-cap: "Scatterplot of Manufacturing Share of GDP versus DN."

# Ensure data is cleaned (remove NA values for manufacturingsharegdp and dn)
analysis_cleaned <- analysis_data %>%
  filter(!is.na(manufacturingsharegdp) & !is.na(dn))

# Create the scatterplot
ggplot(analysis_cleaned, aes(x = manufacturingsharegdp, y = log(dn))) +
  geom_point(alpha = 0.7, color = "steelblue", size = 0.5) +
  labs(
    x = "Manufacturing Share of GDP (%)",
    y = "Log of DN"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

# Linear Models

In this section, we present four models to assess the relationship between Night-Time Light (NTL) intensity and various economic indicators, with a particular focus on understanding how NTL correlates with GDP and how this relationship varies across countries with differing levels of statistical capacity. A key challenge in our analysis was the high correlation between GDP, manufacturing share of GDP, and population, as highlighted in the correlation matrix provided in the appendix. This multicollinearity made it impractical to include these variables in the same model without risking inflated standard errors and unreliable coefficients. To address this, we developed separate models for each variable.

The first model examines the relationship between NTL intensity and GDP, incorporating country-specific and year-specific fixed effects. The inclusion of country-specific fixed effects accounts for unobserved factors that differ across countries, such as geographic characteristics, governance structures, or cultural differences. Year-specific fixed effects are included to capture variations over time, such as global economic trends or technological advancements, that may affect the relationship between NTL and GDP.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \log(GDP_{i,t}) + \alpha_i + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - $alpha_i$ Country fixed effects
  - $gamma_t$ Year fixed effects

Building on the first model, the second model investigates how the relationship between NTL and GDP varies across countries with differing statistical capacities. As seen in Figure 1, the scatterplot of GDP and NTL faceted by statistical capacity suggests a weaker correlation in countries with lower statistical capacity. To explore this further, we include an interaction term between GDP and statistical capacity grade, allowing us to examine whether countries with stronger statistical systems exhibit a more consistent and robust relationship between NTL and GDP. The intuition is that higher statistical capacity is associated with better data quality and more stable economic structures, which may enhance the reliability of NTL as a proxy for economic activity. Similar to the first model, we control for year-specific and country-specific fixed effects to account for temporal and cross-country heterogeneity.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \log(GDP_{i,t}) + \beta_2 (\text{Grade}_{i}) + \beta_3 (\log(GDP_{i,t}) \times \text{Grade}_{i}) + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - Interaction term: $$ \log(\text{GDP}) \times \text{Grade} $$
  - $gamma_t$: Year fixed effects

The third model examines the relationship between NTL intensity and population, while controlling for country and year fixed effects. Population is a key driver of light emissions, particularly in urban areas, and this model isolates its unique contribution to NTL by excluding GDP to address multicollinearity. This approach helps assess whether NTL can serve as a proxy for population, especially in regions where population data may be less reliable, while still accounting for variations across countries and years.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \log(Population_{i,t}) + \alpha_i + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - $alpha_i$: Country fixed effects
  - $gamma_t$: Year fixed effects

The fourth model examines the relationship between NTL intensity and the manufacturing share of GDP, controlling for country and year fixed effects. Manufacturing activity, often concentrated in industrial areas, significantly influences light emissions. By isolating this variable, the model highlights how industrialization shapes NTL patterns, independent of overall GDP, while accounting for country- and time-specific factors.

$$
\log(DN_{i,t}) = \beta_0 + \beta_1 \text{Manufacturing Share}_{i,t} + \alpha_i + \gamma_t + \epsilon_{i,t}
$$

- Where:
  - $alpha_i$: Country fixed effects
  - $gamma_t$: Year fixed effects

# Model Summary

```{r}

#| echo: false
#| warning: false
#| message: false
#| label: tbl-summodel
#| fig-cap: "Model Summary"

model1 <- readRDS(here("models/model1.rds"))
model2 <- readRDS(here("models/model2.rds"))
model3 <- readRDS(here("models/model3.rds"))
model4 <- readRDS(here("models/model4.rds"))
# List of your models
models <- list(
  model1,
  model2,
  model3,
  model4
)

# Function to clean model output by removing fixed effects
clean_model <- function(model) {
  tidy(model) %>%
    filter(!grepl("factor\\(country\\)|factor\\(year\\)", term)) %>% # Remove fixed effects coefficients
    rename(
      Term = term,
      Estimate = estimate,
      `Std. Error` = std.error,
      `t-value` = statistic,
      `p-value` = p.value
    )
}

# Generate a cleaned version of the model summaries
cleaned_models <- lapply(models, clean_model)

# Create a modelsummary table, keeping AIC and BIC
model_summary_table <- modelsummary(
  models,
  coef_omit = "factor\\(country\\)|factor\\(year\\)", # Exclude fixed effects
  gof_map = c("r.squared", "adj.r.squared", "AIC", "BIC", "nobs") # Include key statistics
)

# Display the table
model_summary_table
```

# Appendix

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tab-grades-countries-latex
#| tbl-cap: "List of Countries Grouped by Grade."

# Group countries by grade and collapse into a single string
grades_latex_table <- analysis_data %>%
  select(country, grade) %>%
  distinct() %>%
  group_by(grade) %>%
  summarize(countries = paste(country, collapse = ", ")) %>%
  arrange(grade)

# Generate LaTeX code for the table
grades_latex_table %>%
  knitr::kable(
    format = "latex",
    align = c("l", "p{10cm}"), # Use a wide column for the country list
    col.names = c("Grade", "Countries"),
    escape = FALSE             # Avoid escaping LaTeX special characters
  ) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("striped", "hold_position")
  )

```

```{r fig.width=12, fig.height=9}
#| echo: false
#| warning: false
#| message: false
#| label: fig-man
#| fig-cap: "Manufacturing share of GDP for countries in 2020, showing significant variation across nations, with the highest shares observed in Puerto Rico, Liechtenstein, and Ireland. Countries with higher manufacturing shares are expected to exhibit brighter night-time light intensity due to industrial activity."

# Filter data for the year 2020 with manufacturing share of GDP >= 10%
filtered_data_2020_high_manufacturing <- analysis_data %>%
  filter(year == 2020, manufacturingsharegdp >= 10)

# Create a bar plot for countries with manufacturing share >= 10%
ggplot(filtered_data_2020_high_manufacturing, aes(x = reorder(country, manufacturingsharegdp), y = manufacturingsharegdp)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  labs(
    x = "Country",
    y = "Manufacturing Share of GDP (%)"
  ) +
  coord_flip() +  # Flip coordinates for better readability
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.y = element_text(size = 8)
  )

```

```{r fig.width=12, fig.height=9}
#| echo: false
#| warning: false
#| message: false
#| label: fig-man2
#| fig-cap: "Manufacturing share of GDP for countries in 1992, showing significant variation across nations, with the highest shares observed in Puerto Rico, Ukrain, and Belarus."
# Filter data for the year 2020 with manufacturing share of GDP >= 10%
filtered_data_2020_high_manufacturing <- analysis_data %>%
  filter(year == 1992, manufacturingsharegdp >= 10)

# Create a bar plot for countries with manufacturing share >= 10%
ggplot(filtered_data_2020_high_manufacturing, aes(x = reorder(country, manufacturingsharegdp), y = manufacturingsharegdp)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  labs(
    x = "Country",
    y = "Manufacturing Share of GDP (%)"
  ) +
  coord_flip() +  # Flip coordinates for better readability
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.y = element_text(size = 8)
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: scatterplot-matrix
#| fig-cap: "Scatterplot matrix showing pairwise relationships between key variables in the analysis_data dataset."

# Load necessary library
library(GGally)
library(ggplot2)

analysis_transformed <- analysis_data %>%
  mutate(
    log_dn = log(dn),
    log_gdp = log(gdp),
    log_population = log(population)
  ) %>%
  filter(complete.cases(.)) %>%
  select(log_dn, log_gdp, manufacturingsharegdp, log_population)

# Create the scatterplot matrix
ggpairs(
  analysis_transformed,
  title = "Scatterplot Matrix of Key Variables",
  lower = list(
    continuous = wrap("points", alpha = 0.4, size = 0.1) # Smaller, semi-transparent points
  ),
  diag = list(
    continuous = wrap("barDiag", color = "white", fill = "gray50", alpha = 0.8) # Cleaner histograms
  ),
  upper = list(
    continuous = wrap("cor", size = 4, alignPercent = 0.8) # Correlation coefficients
  )
) +
  theme_minimal(base_size = 10) + # Apply a minimal theme
  theme(
    panel.grid = element_blank(), # Remove gridlines for a cleaner look
    axis.text = element_text(size = 8), # Adjust axis text size
    axis.title = element_text(size = 10, face = "bold") # Bold axis titles
  )

```

# Reference